---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

```{r, echo=TRUE}
#loading libraries

library(dplyr)
library (ggplot2)

# Unzip and pre-procesing the data

filename <- "activity.zip"

unzip(filename)

activity<-read.csv("activity.csv", header=TRUE)

head(activity)


```




## What is mean total number of steps taken per day?

*For this part of the assignment, you can ignore the missing values in the dataset.*

*1. Calculate the total number of steps taken per day*

*2. If you do not understand the difference between a histogram and a barplot, research the difference between them. Make a histogram of the total number of steps taken each day*

*3. Calculate and report the mean and median of the total number of steps taken per day*

**Total Steps per day**

```{r, echo=TRUE}


# create and print number of steps per day

StepsByDay<-aggregate(steps~date, activity, sum)

StepsByDay

```

**Histogram of the total number of steps taken each day**

```{r, echo=TRUE, fig.width=12}

hist(StepsByDay$steps, xlab="Number of Steps", ylab="Number of Days", main="Total Number of Steps by day")

```

**3. Calculate and report the mean and median of the total number of steps taken per day**
```{r, echo=TRUE}
# Mean
meansteps<-mean(StepsByDay$steps)
meansteps

#Median
medianSteps<-median(StepsByDay$steps)
medianSteps

```



## What is the average daily activity pattern?

**1. Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)**   


```{r, echo=TRUE}

steps_by_Interval<-aggregate(steps~interval, data=activity, mean, na.rm=TRUE)

g<-ggplot(steps_by_Interval, aes(interval, steps))

g+geom_line(col="blue")+ggtitle("Average steps per time interval")+xlab("Time")+ylab("Steps")+theme(plot.title = element_text(face="bold", size=18))

```

**2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?**

```{r, echo=TRUE}

five_int_by_max_Steps <- steps_by_Interval[which.max(steps_by_Interval$steps),]$interval
five_int_by_max_Steps


```




## Imputing missing values





## Are there differences in activity patterns between weekdays and weekends?
